---
title: "Assignment3_AMOD_5240H"
author: "Rachael Joan Dias"
date: "02/11/2018"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1 (4.2):

```{r}
#Upper and lower limit of confidence interval
UL<-77
LL<-65

#Sample Mean
(xbar<-(UL+LL)/2)

#Margin of Error
(ME<-(UL-LL)/2)

#degrees of freedom
n<-25
df<-n-1

#t-value
t_value<-qt(.95,df)

#Standard Deviation
(sd<-ME*sqrt(n)/t_value)
```

## Question 2 (4.4):

(a) Since $H_A : \mu>\mu_0$ we do a right tailed test, we get a $P$-value of 0.04260244 which is smaller than $\alpha$=0.05, hence we reject $H_0$.
```{r}
#Part a
n<-11
t<-1.91

#p-value
1-pt(t,n-1)
```
(b) Since $H_A : \mu<\mu_0$ we do a left tailed test, we get a $P$-value of 0.001646786 which is smaller than $\alpha$=0.05, hence we reject $H_0$.
```{r}
#Part b
n<-17
t<--3.45

#p-value
pt(t,n-1)
```

(c) Since $H_A : \mu\neq\mu_0$ we do a two tailed test, we get a $P$-value of 0.4383084 is quite large as compared to $\alpha$=0.05, we fail to reject $H_0$.
```{r}
#Part c
n<-7
t<-0.83

#p-value
2*(1-pt(t,n-1))
```
(d) Since $H_A : \mu>\mu_0$ we do a right tailed test, we get a $P$-value of 0.02121769 which is smaller than $\alpha$=0.05, we reject $H_0$.
```{r}
#Part d
n<-28
t<-2.13

#p-value
1-pt(t,n-1)
```

## Question 3 (4.30):

Hypothesis: 

$H_0 :\mu_{Automatic}=\mu_{Manual}$ (There is no difference between the mean fuel efficiency of cars with manual and those with automatic transmissions.)

$H_A :\mu_{Automatic}\neq\mu_{Manual}$ (There is a significant difference between the mean fuel efficiency of cars with manual and those with automatic transmissions.)

```{r}
#Mean
xbar_auto<-16.12
xbar_man<-19.85

#Standard deviation
sd_auto<-3.58
sd_man<-4.51

#Sample size
n_auto<-26
n_man<-26

#SE
SE<-sqrt((sd_auto^2)/n_auto+(sd_man^2)/n_man)

#t-statistic
(t<-(xbar_auto-xbar_man)/SE)

#degrees of freedom
df<-25

#p-value
2*(pt(t,df))
```

Conclusion: A $P$-value of 0.002883615 is much less than $\alpha$=0.05, hence we reject $H_0$. In the context of this problem we can conclude that the mean fuel efficiency of cars with manual transmission is significantly different from those with automatic transmission.

## Question 4 (4.40):

(a) Hypothesis:

$H_0 :\mu_{Less than HS}=\mu_{HS}=\mu_{Jr Coll}=\mu_{Bachelor's}=\mu_{Graduate}$ The average number of hours worked per week is the same for all the 5 groups.

$H_A :$ The average number of hours worked per week is not the same for all the 5 groups, at least one means is different from the others.

(b) Conditions and assumptions:

1.) Independence between groups: We assume that the samples were randomly collected from a population.

2.) Independence within groups: We assume that the samples were properly taken.

3.) From the boxplot we can see that there is not much difference between the variance of the hours worked per week for the 5 groups. The variance is almost the same for the 5 groups.

4.) Since the samples collected from each group are greater than 30 the distributions of the average working hours per week are approximately normal.

(c) Table values
```{r}
#degrees of freedom
k<-5
n<-1172

#degrees of freedom between
(df_between<-k-1)

#degrees of freedom within
(df_within<-n-k)

#Total
n-1

#Sum of Squares between
MS_between<-501.54
(SS_between<-(k-1)*MS_between)

#Sum of Squares total
SS_within<-267382
(SS_total<-SS_between+SS_within)

#Mean Square within
(MS_within<-SS_within/(df_within))

#F-value
(f_value<-MS_between/MS_within)
```

(d) Conclusion: Assume $\alpha$=0.05, since we get a $P$-value of 0.0682 which is greater than $\alpha$=0.05, we fail to reject $H_0$. In the context of this problem we can conclude that there is no significant difference between the average hours worked per week for the 5 different groups.

## Question 5:

(a)
```{r}
v<-c(0.881,0.863,0.775,0.854,0.810,0.858,0.818,0.768,0.803,0.833,0.742,0.832,0.807,0.841,0.932)

#Mean
(xbar<-mean(v))

#Standard deviation
(s<-sd(v))

#No. of observations
(n<-length(v))
```

(b) Discussion: As the qqplot is close to being a straight line, we can assume normality. The histogram also shows an almost normal distribution, even though the sample size is small.
```{r}
#Histogram
hist(v, col = "blue", xlab="Weight of Blue M&Ms", main="Histogram of M&Ms weight")

#qqplot
qqnorm(v, ylab="Weight of M&Ms", main="Normal Probability Plot")
qqline(v, col="blue")
```



(c) Hypotheses:

$H_0 :\mu=.82$ (the mean weight of the Blue M&Ms is equal to .82)

$H_A :\mu>.82$ (the mean weight of the Blue M&Ms is greater than .82 )

```{r}
#Test-statistic
mu<-.82
(t<-(xbar-mu)/(s/sqrt(n)))

#p-value
1-pt(t,n-1)
```
Conclusion: Since we get a $P$-value of 0.2686079 which is greater than $\alpha$=0.05, we fail to reject $H_0$. In the context of this problem, we can conclude that the mean weight of the Blue M&Ms is not greater than .82.

(d)

```{r}
xvals<-seq(-4,4,.01)
yvals<-dt(xvals,14)
plot(xvals,yvals, type="l", lty=1,ylab="Probability Distribution",xlab="t-statistic",col="black") 
abline(v=0.6325605,lty=2, col="blue")
xvals.sub<- c(0.6325605,  xvals[xvals > 0.6325605],4)
yvals.sub<- c(0, yvals[xvals > 0.6325605], 0)
polygon(xvals.sub,yvals.sub, col="red")
arrows(2,0.15,1,.05)
text(3,.17,labels="area of 0.2686079")
title("T-density Plot, 14df")
```

## Question 6:

```{r}
#Save the treatment and control groups as vectors
control<-c(46, 22, 10, 14, 26, 29, 29, 47, 20, 13)
treatment<-c(2, 32, 30, 3, 14, 32, 2, 39, 18, 2)

#Difference between treatment and control groups
diff<-treatment-control
```

Hypotheses:

$H_0 :\mu_{diff}=0$ (the mean difference between the treatment and control group is 0, i.e no difference)

$H_A :\mu_{diff}<0$ (there is a difference between the treatment and control groups, the soarness levels is less for the treatment group)

The qqplot almost forms a straight line, hence we can justify the assumption of normality.
```{r}
##Histogram and qqplot
qqnorm(diff, ylab="Diff b/w Treatment and Control Groups", main="Normal Probability Plot")
qqline(diff, col="blue")
hist(diff, xlab="Difference between Treatment and Control Groups", col="cyan",
main = "Histogram of Difference")
```

```{r}
#Test Statistic and p-value
xbar<-mean(diff)
s<-sd(diff)
n<-length(diff)
SE<-s/sqrt(n)
(tstat<-xbar/SE)
(pt(tstat,n-1))
t.test(diff, alternative="less")
```
Conclusion: Since we get a $P$-value of 0.09312 which is greater than $\alpha$=0.01, we fail to reject $H_0$. In the context of this problem, we can conclude the treatment did not reduce the soarness in the arm significantly.

## Question 7:

(a) The mean of those who were assigned 10% is 15.12982, while the mean of those who were assigned 65% is 39.50746. The sample means of those who were assigned 10% is much lesser than those who were assigned 65%

```{r}
#Read csv file and save it as a dataframe
Africa<-read.csv(file="anchoring.csv", header=TRUE, stringsAsFactors = FALSE)

#Filter the students who were assigned 10% and save their estimate in Africa_10
Africa_10<-Africa[Africa$Starting.point==10,"Africa.Estimate"]

#Filter the students who were assigned 65% and save their estimate in Africa_65
Africa_65<-Africa[Africa$Starting.point==65,"Africa.Estimate"]

#Means
(xbar_10<-mean(Africa_10))
(xbar_65<-mean(Africa_65))
```
(b) 

Assumptions: 

1. The students were randomly assigned 10% and 65% starting points.

2. The 2 samples do not have any relationship with each other, they are independant of each other. Both the sample sizes are greater than 30.

Hypothesis:

$H_0 :\mu_{10}=\mu_{65}$ (the mean estimate of United Nation member states in Africa is the same for students who were assigned 10% and 65%)

$H_A :\mu_{10}<\mu_{65}$ (the mean estimate of United Nation member states in Africa is lower for those who were assigned 10% than those who were assigned 65%)
```{r}
#t-statistic and P-value
sd_10<-sd(Africa_10)
sd_65<-sd(Africa_65)
n_10<-length(Africa_10)
n_65<-length(Africa_65)


SE<-sqrt(sd_10^2/n_10 + xbar_65^2/sd_65) 
(tstat<-(xbar_10-xbar_65)/SE)
Afric_df<-min(n_10-1,n_65-1)
(pt(tstat,Afric_df))
```

Conclusion: Since we get a $P$-value of 0.003816838 which is less than $\alpha$=0.05, we reject $H_0$. In the context of this problem, we can conclude that the students who received a quantity of 10% have a lower mean estimate of United Nations member states in Africa as compared to the group who were assigned 65%. Thus, we can also conclude that a starting point for a numerical answer can be a source of bias in surveys.


## Question 8:

(a)
```{r}
#Read csv file and save it as a dataframe
cars<-read.csv(file="chest.csv", header=TRUE, stringsAsFactors = FALSE)
```

(b) The standard deviation variations are not too high, as the largest standard deviation is less than twice the smallest standard deviation. It fulfills the standard deviation assumption.
```{r}
#Mean weight for 3 groups
(cars_mean<-tapply(cars$CHEST,INDEX=cars$SIZE,FUN=mean))

#Standard Deviation for 3 groups
(cars_sd<-tapply(cars$CHEST,INDEX=cars$SIZE,FUN=sd)) 
```
(c) From the boxplot we can see that the means are different for the 3 groups of cars. 
```{r}
#Comparative boxplots with means represented by x
boxplot(cars$CHEST ~ cars$SIZE, col="purple", 
main="Boxplot of Chest Deceleration for 3 different car sizes", xlab="Car Sizes", 
ylab="Chest Deceleration")
points(1:3, cars_mean, pch = 4)  
```

(d) Hypothesis:

$H_0 :\mu_{small}=\mu_{midsize}=\mu_{large}$ (the mean chest deceleration measurements for the 3 groups of cars (Small, Medium and Large) are the same)

$H_A :$ At least one mean of chest deceleration measurements is different among the 3 car groups or at least one mean is not equal to another.

(e)
```{r}
#ANOVA test
cars_anova<-aov(CHEST~SIZE, data = cars)
summary(cars_anova)
```

Conclusion: Since we get a $P$-value of 0.0606  which is greater than $\alpha$=0.01, we fail to reject $H_0$. In the context of this problem, we can conclude that the mean of chest deceleration measurements is not significantly different among the 3 groups of cars or there is no mean which is different from the others.
